Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-07-02 article-models-ttinternal.tex
lots of small changes

\begin{lem}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
Then $\cftctxext[\extalg{A}]$ and $\cftfamext[\extalg{A}]$ 
correctly interpret the rules for context
extension and family extension.

More precisely, for $x_0:\cftalgc{\cftalg{A}}$, 
$x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
$x_2:\subst{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftalgf{\cftalg{A}}}$ 
we have the judgmental equality
\begin{align*}
\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftctxext[\extalg{A}]}}
& \jdeq
  \subst
    {x_2}
    {{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftctxext[\extalg{A}]}}.
  \intertext{%
and for $y_0:\cftalgf{\extalg{A}}$, 
$y_1:\subst{y_0}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}$ and 
$y_2:
  \subst
    {y_1}
    { {y_0}
      { \jcomp
          {}
          {\cftfamext[\extalg{A}]}
          {\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}
        }
      }
$
we have the judgmental equality}
\subst
  {{y_2}{{y_1}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftfamext[\extalg{A}]}}}}
  {{y_0}{\cftfamext[\extalg{A}]}}
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
\end{align*}
\end{lem}

\begin{proof}
Both proofs are simple calculations. For the first judgmental equality we have
\begin{align*}
\subst{{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}{{x_0}{\cftctxext[\cftalg{A}]}}
& \jdeq
  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{\tmext{\subst{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{x_2}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{x_2}{{{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{\cftctxext[\cftalg{A}]}}.
\end{align*}
and to prove the second judgmental equality we calculate
\begin{align*}
\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}{{y_0}{\cftfamext[\cftalg{A}]}}
& \jdeq
  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}
        }
      }
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}{\cftfamext[\cftalg{A}]}}
        }
      }
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{\tmext{\subst{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{y_2}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}.\qedhere
\end{align*}
\end{proof}

@@

\begin{comment}
\begin{rmk}
I suspect that if we copy this theory of extension algebras to Martin-L\"of
type theory, with the judgmental equalities replaced by identifications, with
dependent pair types rather than those strict extensions, etcetera, then
the type of $f_1$ for which these two diagrams commute is a mere proposition.

With this notion of morphism, a term of an extension
algebra $(A,P,\epsilon_0,\epsilon_1)$ is a pair $(x_0,x_1)$ such that
$\subst{x_1}{{x_0}{\epsilon_0}}\jdeq x_0$.
\end{rmk}
\end{comment}

\begin{comment}
Extension algebras don't come in isolation. There are also extension algebra
families and extension algebra terms. We now aim to define these and to
establish various constructions of new extension algebras out of old ones:
the empty extension algebra, and extensions, weakenings and substitutions
of extension algebras and of course the identity term as an extension algebra
term. We start with extension algebra families.

\begin{defn}
Consider an extension algebra $\mathcal{A}\defeq(A,P,e,f)$. 
An extension algebra family over $\mathcal{A}$ is likewise a quadruple
$\mathcal{B}\defeq(B,Q,g,h)$. Here we have a family $\jfam{{\Gamma}{A}}{B}$, a
family $\jfam{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{Q}$ and
\begin{align*}
\jhom*{{{\Gamma}{A}}{P}}{\ctxext{\ctxwk{P}{B}}{Q}}{\jcomp{}{\epsilon_0}{B}}{g}\\
\jhom*{{{{\Gamma}{A}}{P}}{\ctxwk{P}{B}}}{\ctxext{Q}{\jcomp{}{g}{Q}}}{Q}{h}.
\end{align*}
The quadruple $(\jcomp{}{\epsilon_0}{B},Q,g,h)$ is required to be an extension algebra
in context $\ctxext{{\Gamma}{A}}{P}$.
\end{defn}

\begin{defn}
Suppose $\mathcal{A}$ is an extension algebra and $\mathcal{B}$ is an extension
algebra family over $\mathcal{A}$. A term of $\mathcal{B}$ is a pair $(x,y)$
consisting of
\begin{align*}
\jterm*{{\Gamma}{A}}{B}{x}\\
\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{Q}}{y}
\end{align*}
such that the diagrams
\begin{equation*}
\begin{tikzcd}
\ctxext{\jcomp{}{\epsilon_0}{B}}{Q} 
  \ar{r}{g} 
  \ar[shift right=.7ex,fib]{d}
& B 
  \ar[shift right=.7ex,fib]{d} 
  \\
\ctxext{A}{P} 
  \ar[shift right=.7ex,dotted]{u}[swap]{\tmext{\jcomp{}{\epsilon_0}{x}}{y}}
  \ar{r}{e}
& A
  \ar[shift right=.7ex,dotted]{u}[swap]{x}
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}
\jcomp{}{f}{\ctxext{Q}{\jcomp{}{g}{Q}}}
  \ar{r}{\jcomp{}{f}{h}}
  \ar[shift right=.7ex,fib]{d}
& Q
  \ar[shift right=.7ex,fib]{d}
  \\
\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}
  \ar{r}{\idtm{\jcomp{}{f}{\jcomp{}{\epsilon_0}{B}}}}
  \ar[shift right=.7ex,fib]{d}
  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{\jcomp{}{f}{y}}
& \jcomp{}{\epsilon_0}{B}
  \ar[shift right=.7ex,fib]{d}
  \ar[shift right=.7ex,dotted,mapsto]{u}[swap]{y}
  \\
\ctxext{P}{\jcomp{}{\epsilon_0}{P}}
  \ar{r}[swap]{f}
  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{f}{\jcomp{}{\epsilon_0}{x}}}
& P
  \ar[shift right=.7ex,dotted]{u}[swap]{\jcomp{}{\epsilon_0}{x}}
\end{tikzcd}
\end{equation*}
commute.
\end{defn}

\begin{defn}
Suppose $\mathcal{A}$ and $\mathcal{B}$ are extension algebras in context
$\Gamma$. We define the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$
to be the quadruple
\begin{equation*}
(\ctxwk{A}{B},\ctxwk{\ctxext{A}{P}}{Q},\ctxwk{\ctxext{A}{P}}{g},\ctxwk{\ctxext{A}{P}}{h}).
\end{equation*}
Note that $\ctxwk{\ctxext{A}{P}}{Q}$ is a family over $\ctxwk{\ctxext{A}{P}}{B}$,
whereas it should be a family over $\jcomp{}{\epsilon_0}{\ctxwk{A}{B}}$. These are the
same by \autoref{lem:prehom}.
\end{defn}

\begin{rmk}
Before we continue, let us explore what it means to be an extension algebra
term of the extension algebra $\ctxwk{\mathcal{A}}{\mathcal{B}}$. Such an
extension algebra term $(x,y)$ would consist of
\begin{align*}
\jterm*{{\Gamma}{A}}{\ctxwk{A}{B}}{x}\\
\jterm*{{{\Gamma}{A}}{P}}{\subst{\jcomp{}{\epsilon_0}{x}}{\ctxwk{\ctxext{A}{P}}{Q}}}{y}.
\end{align*}
Thus, $x$ is a context morphism from $A$ to $B$ and $y$ is nothing but a term
of $\jcomp{}{\jcomp{}{\epsilon_0}{x}}{Q}$. For $x$, we see that the diagram
\begin{equation*}
\begin{tikzcd}
\ctxext{B}{Q} 
  \ar{r}{g} 
& B 
  \\
\ctxext{A}{P} 
  \ar{u}{\jvcomp{}{x}{y}}
  \ar{r}{e}
& A
  \ar{u}[swap]{x}
\end{tikzcd}
\end{equation*}
commutes.
\end{rmk}
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CFT-algebras}
The notion of CFT-algebras that we will
study in this subsection will also have an empty object $(\cftempc{\cftalg{A}},
\cftempf{\cftalg{A}})$ and guarantiees that families over the empty object
are just the terms of the underlying family $\cftalgc{\cftalg{A}}$ of contexts
of the term-algebra $\cftalg{A}$. We will begin with an auxilary lemma which
will help us dealing with the internal empty contexts:

@@

\begin{comment}
\begin{defn}
A quintuple $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ consisting of
\begin{enumerate}
\item A family $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
\item A family $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
\item A family $\cftalgt{\cftalg{A}}$ in context $\ctxext{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}$,
\item A term $\cftempc{\cftalg{A}}$ of $\cftalgc{\cftalg{A}}$ in context $\Gamma$,
\item A term $\cftempf{\cftalg{A}}$ of $\cftalgf{\cftalg{A}}$ in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
\end{enumerate}
is said to be an \emph{empty-object algebra in context $\Gamma$} if additionally
the following judgmental equalities hold:
\begin{align}
\jfameq*{\Gamma}{\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}{\cftalgc{\cftalg{A}}}
  \label{empalg-eq1}
  \\
\jtermeq*{\Gamma}{\cftalgc{\cftalg{A}}}{\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}{\cftempc{\cftalg{A}}}.
  \label{empalg-eq2}
%  \\
%\jfameq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftalgt{\cftalg{A}}}}{\emptyf}
\end{align}
When $(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ is an empty-object algebra
in context $\Gamma$, we will
also refer to the pair $(\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$ as the
\emph{empty object of $\cftalg{A}$}.
\end{defn}

Below, we give the definition of empty-object homomorphisms. They have, like
extension homomorphisms, an action on contexts and an action on families and
in addition they have an action on terms. Thus, we can picture a term
homomorphism as follows:
\begin{equation*}
\begin{tikzcd}
\cftalgt{\cftalg{A}}
  \ar[fib]{d}
  \ar{r}{\cfthomt{\cfthom{f}}}
& \cftalgt{\cftalg{B}}
  \ar[fib]{d}
  \\
\cftalgf{\cftalg{A}}
  \ar[fib]{d}
  \ar{r}{\cfthomf{\cfthom{f}}}
& \cftalgf{\cftalg{B}}
  \ar[fib]{d}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
& \cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation*}

\begin{defn}
Let $\cftalg{A}$ and $\cftalg{B}$ be term algebras in context $\Gamma$.
An \emph{empty-object homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$} is a triple 
$(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$ consisting of
\begin{align*}
\jhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}\\
\jfhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}\\
\jfhom*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
satisfying the judgmental equalities
\begin{align*}
\jtermeq*
  {\Gamma}
  {\cftalgc{\cftalg{B}}}
  {\subst{\cftempc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}}
  {\cftempc{\cftalg{B}}}
  \\
\jhomeq*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\subst{\cftempc{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
  {\cfthomc{\cfthom{f}}}
  \\
\jtermeq*
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{f}}}}
  {\jcomp{\cftalgc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cftempf{\cftalg{B}}}}.
\end{align*}
\end{defn}

A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
\begin{equation*}
(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}],\cftempc{\cftalg{A}},
\cftempf{\cftalg{A}})
\end{equation*}
for which 
the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty-object algebra in context
$\Gamma$, satisfying additional judgmental equalities expressing the 
compatibility of $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$ with 
$\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
There will be five such judgmental equalities. Four of them will express that
extension acts as the identity term when applied to the empty object in one of
its arguments. The other expresses that family extension restricted to
families above the empty object is the same as context extension. To be able
to state these, we need to know that $\unfold{\empalg{A}}$ is an empty-object
algebra. In the following, we verify step-by-step that we can indeed make the
suggested requirements.
\begin{enumerate}
\item The type of $\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}$ is
$\subst
  { \cftempc{\cftalg{A}}
    }
  { \ctxwk
      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
      {\cftalgc{\cftalg{A}}}
    }$.
We have the judgmental equalities
\begin{align*}
\subst
  { \cftempc{\cftalg{A}}
    }
  { \ctxwk
      {\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
      {\cftalgc{\cftalg{A}}}
    }
& \jdeq
  \subst
    { \cftempc{\cftalg{A}}
      }
    { \ctxwk
        { \cftalgf{\cftalg{A}}
          }
        { {\cftalgc{\cftalg{A}}}
          {\cftalgc{\cftalg{A}}}
          }
      }
  \tag{by \autoref{comp-ew-f}}
  \\
& \jdeq
  \ctxwk
    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
      }
    { \subst{\cftempc{\cftalg{A}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
      }
  \tag{by \autoref{comp-sw-f}}
  \\
& \jdeq
  \ctxwk
    { \subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
      }
    { A
      }
  \tag{by \autoref{cancellation-ws-f}}
  \\
& \jdeq
  \ctxwk{A}{A}
  \tag{by \autoref{empalg-eq2}}
\end{align*}
Thus, we require that
\begin{equation}
\jtermeq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq1}
\end{equation}
\end{enumerate}

We can immediately state the first two:
\begin{align}
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq1}
  \\
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq2}
\end{align}
To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
we use \autoref{lem:empalg-mor}. Note that in combination with 
\autoref{cftalg-eq1} and \autoref{cftalg-eq2} we get the judgmental
equalities
\begin{equation}\label{cor:empalg-mor}
\jfameq{{\Gamma}{A}}{\subst{\phi_i}{\jcomp{}{\epsilon_0}{P}}}{P}
\end{equation}
for $i$ being $\mathsf{c}$ or $\mathsf{f}$. Consequently, 
$\subst{\phi_0}{\epsilon_1}$ is a morphism from $\ctxext{A}{P}$ to $A$, so we 
can require that
\begin{equation}\label{cftalg-eq5}
\jhomeq{\Gamma}{{A}{P}}{A}{\subst{\phi_0}{\epsilon_1}}{\epsilon_0}.
\end{equation}
This judgmental equality expresses that family extension restricted to the empty object
$\phi_0$ is context extension. We also have two judgmental
equalities expressing that $\epsilon_1$ restricted to $\phi_1$ and
$\jcomp{}{\epsilon_0}{\phi_1}$ are both the identity term on $P$. Note that
the term $\jcomp{}{\epsilon_0}{\phi_1}$ serves as the empty family of families,
being a term of $\jcomp{}{\epsilon_0}{P}$.

Before we can require that $\subst{\phi_1}{\epsilon_1}\jdeq\idtm{P}$, we need
to compute the family
\begin{equation*}
\jfam
  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
  {\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
\end{equation*}
of which $\subst{\phi_1}{\epsilon_1}$ is a term.
We have already established that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq
P$. We also see that 
\begin{align*}
\subst{\phi_1}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
& \jdeq
  \subst{\phi_1}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
  \tag{by \autoref{comp-ew-f}}
  \\
& \jdeq
  \ctxwk{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}{\subst{\phi_1}{\ctxwk{P}{P}}}
  \tag{by \autoref{comp-sw-f}}
  \\
& \jdeq
  \ctxwk{P}{\subst{\phi_1}{\ctxwk{P}{P}}}
  \tag{by \autoref{cor:empalg-mor}}
  \\
& \jdeq
  \ctxwk{P}{P},
  \tag{by \autoref{cancellation-ws-f}}
\end{align*}
so we see that it indeed makes sense to require that
\begin{equation}\label{cftalg-eq3}
\jtermeq{{{\Gamma}{A}}{P}}{\ctxwk{P}{P}}{\subst{\phi_1}{\epsilon_1}}{\idtm{P}}.
\end{equation}
For the last judgmental equality, we need to compute the family
\begin{equation*}
\jfam
  {{{\Gamma}{A}}{\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}}}
  {\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}}.
\end{equation*}
This is easy:
\begin{align*}
\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}{P}}
& \jdeq
  \subst{\jcomp{}{\epsilon_0}{\phi_1}}{\ctxwk{\jcomp{}{\epsilon_0}{P}}{{P}{P}}}
  \\
& \jdeq
  \ctxwk{P}{P}.
\end{align*}
Thus we are allowed to require that
\begin{equation}\label{cftalg-eq4}
\subst{\jcomp{}{\epsilon_0}{\phi_1}}{\epsilon_1}.
\end{equation}

We bring all this together in the definition of extension-empty algebras:

\begin{defn}
An \emph{extension-empty algebra in context $\Gamma$} 
is a sextuple $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ for which 
the quadruple $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context 
$\Gamma$, the quadruple $(A,P,\phi_0,\phi_1)$ is an empty algebra in context
$\Gamma$, satisfying the judgmental equalities 
\autoref{cftalg-eq1,cftalg-eq2,cftalg-eq5,cftalg-eq3,cftalg-eq4}.
\end{defn}
\end{comment}

@@

\subsection{CFT-homomorphisms}
\begin{defn}
A homomorphism of term algebras $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
consists of a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
such that $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is an extension homomorphism
\end{defn}

\subsection{Weakening algebras}
Weakening algebras will be term-algebras with certain added structure.
Although strictly speaking one could formulate a notion of weakening algebra
which only depends on extension algebras and which omits both empty objects
and terms, we shall not do so here.

Let $\cftalg{A}$ be a term-algebra in context $\Gamma$. Weakening will be a
CFT-homomorphism
\begin{equation}\label{cftwk}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\cftalg{F_{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation}
satisfying judgmental equalities which express abstracted analogues of
the rules in sections \autoref{comp-ww,comp-ew,comp-0w}. The
definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.

>>>>

\begin{lem}
For $\gamma:\cftalgc{\cftalg{A}}$, 
$a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and 
$p:\subst{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{\cftalgf{\cftalg{A}}}$ 
we have the judgmental equality
\begin{align*}
\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
& \jdeq
  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
  \intertext{%
and when we also have 
$q: \subst
      {\apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
      {\cftalgf{\extalg{A}}}$, %
we have the judgmental equality}
\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
& \jdeq
  \apply\cftfamext[\extalg{A}]{\gamma}{%
    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.
\end{align*}
\end{lem}

\begin{proof}
Both proofs are simple calculations. For the first judgmental equality we have
\begin{align*}
\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
& \jdeq
  \subst
    {\tmext{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}}
    {\cftctxext[\cftalg{A}]}
  \\
& \jdeq 
  \subst
    { {p}
      { {a}
        { {\gamma}
          {\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}
          }
        }
      }
    { \cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq
  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq 
  \subst{{p}{{a}{{\gamma}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst
    {\tmext{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
    {\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
\end{align*}
and to prove the second judgmental equality we calculate
\begin{align*}
\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
& \jdeq
  \apply\cftfamext[\extalg{A}]{\gamma}{%
    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.\qedhere
\end{align*}
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CFT-algebras}
The notion of CFT-algebras that we will
study in this subsection will be extension algebras which also have an empty 
context $\cftempc{\cftalg{A}}$ and an empty family $\cftempf{\cftalg{A}}$ 
satisfying (among other conditions) that families over the empty context in
$\cftalg{A}$ are just the contexts of $\cftalg{A}$. We will begin with an 
auxilary lemma which will help us dealing with the internal empty contexts:

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CFT-homomorphisms}
\begin{defn}
A homomorphism of term algebras $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
consists of a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
such that $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is an extension homomorphism
\end{defn}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weakening algebras}
Weakening algebras will be CFT-algebras with certain added structure.
Although strictly speaking one could formulate a notion of weakening algebra
which only depends on extension algebras and which omits both empty objects
and terms, we shall not do so here.

Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Weakening will be a
CFT-homomorphism
\begin{equation}\label{cftwk}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\cftalg{F_{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation}
satisfying judgmental equalities which express abstracted analogues of
the rules in sections \autoref{comp-ww,comp-ew,comp-0w}. The
definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.
